\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}



\title{Stance Detection for the Fake News Challenge Dataset using Deep Learning}

\author{Anak Agung Ngurah Bagus Trihatmaja and Shishir Kurhade}

\date{}

\begin{document}

\maketitle

\abstract
The problem of fake news has arisen recently as a threat to high-quality 
journalism and well-informed public disclosure. 
The goal of fake news challenge is to explore how artificial intelligence 
technologies, particularly machine learning and 
natural language processing, might be leveraged to combat the fake news problem.
A dataset consisting of 49,972 news article samples which have been labeled 
into four categories: 
“Unrelated”, “Agrees”, “Disagrees”, and “Discusses” has been provided.
Our goal is to develop machine learning models to predict these labels. 
In this time, we will use Gate Recurrent Unit (GRU) to predict the labels.
It is often said that neural network performs better in this kind of problem,
therefore we will also incorporate logistic regression as our baseline.


\section{Introduction}
Over the past couple of years, the issue of “fake news” \-- defined by the 
New York Times as “made-up stories written with the intention to deceive” and 
published in formats similar to those of traditional “real” news has arisen as 
a threat to high-quality journalism and the society in general. 
In particular,fake news has been accused of increasing political polarization 
and partisan conflict in the United States during the divisive 2016 
presidential campaign and also during the Brexit referendum in 2016.
In our final project for DS5220, we try to combat this serious social problem 
using Machine Learning techniques. This problem views the task of fake-news 
detection as a stance detection problem which is a labelling task, 
where we want to automatically classify a news into four labels, 
which are ‘unrelated’, ‘agrees’, ‘disagrees’, and ‘discusses’.  

A reasoning for these labels is as follows:
\begin{enumerate}
  \item \textbf{Agrees}: The body text agrees with the headline.
  \item \textbf{Disagrees}: The body text disagrees with the headline.
  \item \textbf{Discusses}: The body text discuss the same topic as the headline, 
    but does not take a position
  \item \textbf{Unrelated}: The body text discusses a different topic than the headline
\end{enumerate}
Here, we don’t label the the news as binary (fake news or legitimate news),
because labelling such datasets with binary label tend to be biased towards 
the labeller.The classifier that we build could later be used as a base of a 
fake news detection tool that can automatically categorize the news into the 
stances given.

\section{Proposed Project}
\subsection{Dataset Overview}
The data provided by the Fake News Challenge consists of 
headline, body, and stace. 
Where stace is one of the categories we have mentioned above: 
unrelated, discuss, agree, disagree. 
For training, there are two csv files:
\begin{enumerate}
  \item \textit{train\_bodies.csv}: contains the body text of articles with its ID
  \item \textit{train\_stances.csv}: contains labeled stances for pairs of article 
      headlines and article bodies, in which the article bodies rever to the 
      bodies in train\_bodies.csv
\end{enumerate}
The distribution of the data is as follows:
\begin{center}
  \begin{tabular} 
    {|c|c|c|c|c|}
    \hline
    Rows & Unrelated & Discuss & Agree & Disagree \\
    \hline
    49972 & 0.73131 & 0.17828 & 0.0736012 & 0.0168094 \\
    \hline
  \end{tabular}
\end{center}

For development and training, we will sample our data and 
do data pre-processing.
Roughly, we will use 2000 samples as our development set, 
for choosing hyperparameter and performance evaluation, and leave the rest for 
our training data. 
The data pre-processing includes normalising the case, handling the punctuation
and non-alphabetic symbols.

\subsection{Methods}
\subsubsection{Pre-processing}
The text from the corpus will be converted to tokens using \textit{‘nltk’} package and 
then be mapped to corresponding vectorized forms using  pre-trained \textit{GloVe} 
representations freely available on the Stanford NLP group website. 
Since the text sequences observed will be of variable length we will pad all 
sequences to the length of the maximum length text sequence before inputting 
it to our model.

\subsubsection{Learning Model}
We will used a learning model that is  RNN variant to predict the stances. 
The reason why it is based on RNN instead of CNN is that, for language 
modelling, RNN models are still the best approach (Mikolov et al. 2011).
The difficulty with RNN models is that they are hard to train, because they 
suffer from the vanishing gradient problem (Kevin P. Murphy book page 570). 
To solve the problem is to use RNN variant model called long short-term 
memory (LSTM). 
However, recently a model called gate recurrent unit (GRU) was introduced by 
Cho et al. [2014].
GRU is similar with LSTM. Unlike LSTM, GRU combines forget and input gates 
into a single “update gate” and merges the cell state and hidden state.
This makes GRU computationally more efficient than LSTM and GRU model has 
been increasingly popular. For this project we will use TensorFlow API for GRU,
tf.nn.rnn\_cell.GRUCell. 
It is often said, that neural network performs better for a case like this. 
Therefore, in this chance, we will also perform logistic regression as a 
baseline for our project. We will build the logistic regression from scratch.
To make sure our result is correct, we will use logistic regression provided by
\textit{scikit.learn}.
To measure the accuracy, we will use the evaluation tool provided by 
Fake News Challenge. The tool will evaluate our model and output a score. 
The score is weighted as follows:
\begin{enumerate}
    \item Classifying pair of body and headline as related and unrelated is weighted 25\%
    \item Classifying related pairs as agrees, disagrees, or discusses is weigted as 75\%
\end{enumerate}
The reason behind the weighting is the related/unrelated classification task is expected to be much easier while classifying the agrees, disagrees or discuss is more difficult and more relevant to fake news detection.

\end{document}
